---
title: "markdown"
author: "Drew Dahlquist"
date: "4/8/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Getting data

```{r}
# get all unique player id's
ids = unique(logs$player_id)

# alloc matrix
X = matrix(NA, nrow=length(ids), ncol=2)
rownames(X) = ids
colnames(X) = c("time", "stacks")

# record each player's total time taken and total # of stacks completes
for(id in ids) {
  time = log(max(na.omit(logs[logs$player_id == id,"event_time_dbl"]))) # using log(time)
  stacks = length(unique(na.omit(logs[logs$player_id == id,"stack_id"])))
  X[toString(id),] = c(time,stacks)
}
```

### Raw data plot

```{r}
# plot of data
plot(X,xlab="log(time taken)",ylab="total stacks", pch=20, cex=2)

# median & middle 90th percentile
abline(v=quantile(X[,"time"], c(.25,.5,.75)),col=c("blue","red","blue"))

# print(quantile(exp(X[,"time"]), 0.5)/3600)
# print(mean(exp(X[,"time"]))/3600)
```

### Getting Player1's path

```{r}
# player 1
pid = "6607011"

# alloc matrix
Player1 = matrix(NA, nrow=1641, ncol=2)
colnames(Player1) = c("time", "stacks")

# player data (from logs file)
# player_time = log(na.omit(logs[logs$player_id == pid & !is.na(logs$event_time_dbl) & !is.na(logs$stack_id),"event_time_dbl"])) # using log(time)
# player_stack = na.omit(logs[logs$player_id == pid & !is.na(logs$event_time_dbl) & !is.na(logs$stack_id),"stack_id"])

# player data (from player file)
player_time = log(na.omit(player1[!is.na(player1$event_time_dbl) & !is.na(player1$stack_id),"event_time_dbl"])) # using log(time)
player_stack = na.omit(player1[!is.na(player1$event_time_dbl) & !is.na(player1$stack_id),"stack_id"])

player_stacks = matrix(NA, nrow=1, ncol=length(player_stack))
for(i in 1:length(player_stack)) {
  for(j in 1:i) {
    # num = length(unique(player_stack[1:j]))
    num = length(as.set(player_stack[1:j]))
  }
  player_stacks[,i] = num
}

Player1[,] = c(player_time,player_stacks)
```

### Getting Player2's path

```{r}
# player 2
pid = "6486029"

# alloc matrix
Player2 = matrix(NA, nrow=2170, ncol=2)
colnames(Player2) = c("time", "stacks")

# player data (from logs file)
# player_time = log(na.omit(logs[logs$player_id == pid & !is.na(logs$event_time_dbl) & !is.na(logs$stack_id),"event_time_dbl"])) # using log(time)
# player_stack = na.omit(logs[logs$player_id == pid & !is.na(logs$event_time_dbl) & !is.na(logs$stack_id),"stack_id"])

# player data (from player file)
player_time = log(na.omit(player2[!is.na(player2$event_time_dbl) & !is.na(player2$stack_id),"event_time_dbl"])) # using log(time)
player_stack = na.omit(player2[!is.na(player2$event_time_dbl) & !is.na(player2$stack_id),"stack_id"])

player_stacks = matrix(NA, nrow=1, ncol=length(player_stack))
for(i in 1:length(player_stack)) {
  for(j in 1:i) {
    # num = length(unique(player_stack[1:j]))
    num = length(as.set(player_stack[1:j]))
  }
  player_stacks[,i] = num
}

Player2[,] = c(player_time,player_stacks)
```

### Plotting Player{1,2}'s path on data

```{r}
# plot of data
plot(X,xlab="log(time taken)",ylab="total stacks", pch=20, cex=2)
lines(Player1, type="l", col="blue", lwd=2)
lines(Player2, type="l", col="orange", lwd=2)
legend("topleft", c("Player 1", "Player 2"), lty=c(1,1),col=c("blue","orange"))
```

### K means, I'm not a fan

```{r}
set.seed(1)

km.out = kmeans(X,6,nstart=50)

plot(X[,"time"],X[,"stacks"], col=(km.out$cluster +1), main="K-Means Clustering Results with K=6", xlab="log(time taken)", ylab="total stacks", pch=20, cex=2)
```

6 clusters: 4 (left) of starting, middle, late, & finished players, 2 (right) of players that have given up

### DB scan

```{r}
library(dbscan)

set.seed(1)

db.out = dbscan(X, eps=1.01, minPts=5)

plot(X,col=db.out$cluster+1,main="DBSCAN",xlab="log(time taken)",ylab="total stacks", pch=20, cex=2)

# individual player's paths
# lines(Player1, type="l", col="blue", lwd=2)
# lines(Player2, type="l", col="orange", lwd=2)
# legend("topleft", c("Player 1", "Player 2"), lty=c(1,1),col=c("blue","orange"))
```

Above DB-Scan results in 3 clusters, each of which may be interpreted as different "stages" of players.

Bottom-left has players that are new, but off to good start.

Top-center has players that are making good progress in a reasonable amount of time.

Bottom-right has players that seem to have given up on the game.

NOTE: eps=1.005 gives 4 clusters, splitting up the early-gamers into a small mid-game group.

### standardize time & stacks values to have mean=0, sd=1, then try DBSCAN with that data, since it relies on dist. metric

```{r}
# get all unique player id's
ids = unique(logs$player_id)

# alloc matrix
X = matrix(NA, nrow=length(ids), ncol=2)
rownames(X) = ids
colnames(X) = c("time", "stacks")

# record each player's total time taken and total # of stacks completes
for(id in ids) {
  time = log(max(na.omit(logs[logs$player_id == id,"event_time_dbl"]))) # using log(time)
  stacks = length(unique(na.omit(logs[logs$player_id == id,"stack_id"])))
  X[toString(id),] = c(time,stacks)
}

t_m = mean(X[,"time"])
t_sd = sd(X[,"time"])
s_m = mean(X[,"stacks"])
s_sd = sd(X[,"stacks"])

# alloc matrix for normalized X
X_stand = matrix(NA, nrow=length(ids), ncol=2)
rownames(X_stand) = ids
colnames(X_stand) = c("time", "stacks")

# standardize
for(id in 1:length(ids)) {
  X_stand[id,] = c((X[id,"time"]-t_m)/t_sd,(X[id,"stacks"]-s_m)/s_sd)
}
```

### (X-mu)/sd DBSCAN

```{r}
library(dbscan)

set.seed(1)

db.out = dbscan(X_stand, eps=.5, minPts=5)

plot(X_stand,col=db.out$cluster+1,main="DBSCAN Standardized",xlab="log(time taken)",ylab="total stacks", pch=20, cex=2)
```

### Reachability stuff

```{r}
set.seed(1)

opt.out = optics(X, eps=2, minPts=5)
db = extractDBSCAN(opt.out, eps_cl = 1.01)
plot(db,main="DBSCAN Reachability Plot")

opt.out = optics(X_stand, eps=2, minPts=5)
db = extractDBSCAN(opt.out, eps_cl = 0.5)
plot(db,main="Standardized DBSCAN Reachability Plot")
```

### (X-mu)/sd DBSCAN with players

```{r}
library(dbscan)

set.seed(1)

db.out = dbscan(X_stand, eps=.5, minPts=5)

# alloc matrix for normalized X
Player1_stand = matrix(NA, nrow=1641, ncol=2)
colnames(Player1) = c("time", "stacks")
Player2_stand = matrix(NA, nrow=2170, ncol=2)
colnames(Player1) = c("time", "stacks")

# means & sd's
t_m = mean(X[,"time"])
t_sd = sd(X[,"time"])
s_m = mean(X[,"stacks"])
s_sd = sd(X[,"stacks"])

# standardize
Player1_stand[,] = c((Player1[,"time"]-t_m)/t_sd,(Player1[,"stacks"]-s_m)/s_sd)
Player2_stand[,] = c((Player2[,"time"]-t_m)/t_sd,(Player2[,"stacks"]-s_m)/s_sd)

plot(X_stand,col=db.out$cluster+1,main="Standardized DBSCAN",xlab="log(time taken)",ylab="total stacks", pch=20, cex=2)
lines(Player1_stand, type="l", col="blue", lwd=2)
lines(Player2_stand, type="l", col="orange", lwd=2)
legend("topleft", c("Player 1", "Player 2"), lty=c(1,1),col=c("blue","orange"))
```

### Differences between top two clusters

```{r}
# red = 2 => top-center cluster is # 1
# blue = 4 => top-center cluster is # 3
plot(X_stand,col=1,main="Standardized DBSCAN with Score Data",xlab="log(time taken)",ylab="total stacks", pch=20, cex=2)
points(X_stand[db.out$cluster == 1,],col=2,main="Standardized DBSCAN",xlab="log(time taken)",ylab="total stacks", pch=20, cex=2)
points(X_stand[db.out$cluster == 3,],col=4,main="Standardized DBSCAN",xlab="log(time taken)",ylab="total stacks", pch=20, cex=2)

for(i in 1:150) {
  points(X_stand[toString(scores$player_id[i]),"time"], X_stand[toString(scores$player_id[i]),"stacks"], col=7, pch=1, cex=2)
}
```

```{r}
# red = 2 => top-center cluster is # 1
# blue = 4 => top-center cluster is # 3
plot(X_stand,col=1,main="Standardized DBSCAN with Score Data",xlab="log(time taken)",ylab="total stacks", pch=20, cex=2)
points(X_stand[db.out$cluster == 1,],col=2,main="Standardized DBSCAN",xlab="log(time taken)",ylab="total stacks", pch=20, cex=2)
points(X_stand[db.out$cluster == 3,],col=4,main="Standardized DBSCAN",xlab="log(time taken)",ylab="total stacks", pch=20, cex=2)

cluster1_time_mean = mean(X_stand[db.out$cluster == 1,"time"])
cluster1_stacks_mean = mean(X_stand[db.out$cluster == 1,"stacks"])

cluster3_time_mean = mean(X_stand[db.out$cluster == 3,"time"])
cluster3_stacks_mean = mean(X_stand[db.out$cluster == 3,"stacks"])

points(cluster1_time_mean, cluster1_stacks_mean, col=3, pch=20, cex=3)
points(cluster3_time_mean, cluster3_stacks_mean, col=3, pch=20, cex=3)
```

```{r}
# alloc matrix for mean scores for each id
scores_stats = matrix(NA, nrow=length(unique(na.omit(scores)$player_id)), ncol=4)
rownames(scores_stats) = unique(na.omit(scores)$player_id)
colnames(scores_stats) = c("mean",".25","median",".75")

# compute mean scores
for(id in unique(na.omit(scores)$player_id)) {
  scores_stats[toString(id),] = c(mean(scores[na.omit(scores)$player_id == toString(id),"S5_mean"]),quantile(scores[na.omit(scores)$player_id == toString(id),"S5_mean"], c(0.25,0.50,0.75)))
}

# alloc matrix for cluster stats
cluster1_score_stats = matrix(NA, nrow=length(intersect(rownames(X_stand[db.out$cluster == 1,]),rownames(scores_stats))), ncol=4)
rownames(cluster1_score_stats) = unique(intersect(rownames(X_stand[db.out$cluster == 1,]),rownames(scores_stats)))
colnames(cluster1_score_stats) = c("mean",".25","median",".75")

# cluster 1  scores stats
for(id in intersect(rownames(X_stand[db.out$cluster == 1,]),rownames(scores_stats))) {
  cluster1_score_stats[id,] = scores_stats[id,]
}

# alloc matrix for cluster stats
cluster3_score_stats = matrix(NA, nrow=length(intersect(rownames(X_stand[db.out$cluster == 3,]),rownames(scores_stats))), ncol=4)
rownames(cluster3_score_stats) = unique(intersect(rownames(X_stand[db.out$cluster == 3,]),rownames(scores_stats)))
colnames(cluster3_score_stats) = c("mean",".25","median",".75")

# cluster 3  scores stats
for(id in intersect(rownames(X_stand[db.out$cluster == 3,]),rownames(scores_stats))) {
  cluster3_score_stats[id,] = scores_stats[id,]
}

# median of cluster 1 means, percentiles
# print(median(cluster1_score_stats[,"mean"]))
# print(median(cluster1_score_stats[,".25"]))
# print(median(cluster1_score_stats[,"median"]))
# print(median(cluster1_score_stats[,".75"]))

# print("----------------")

# median of cluster 3 means, percentiles
# print(median(cluster3_score_stats[,"mean"]))
# print(median(cluster3_score_stats[,".25"]))
# print(median(cluster3_score_stats[,"median"]))
# print(median(cluster3_score_stats[,".75"]))

boxplot(cluster1_score_stats[,"mean"], cluster3_score_stats[,"mean"], col=c(2,4), main="Plot of Mean Scores from Clusters 1 and 3")
legend("bottomright", c("Cluster 1", "Cluster 3"), pch=c(20,20), cex=c(2,2), col=c(2,4))
```
